---
title: "R lesson 3"
author: Kelsey Deaton
date: 2018-09-13
output: html_notebook
---

GETTING HELP IN R

```{r}
?"+"

??"+"

?? rnorm()

? rnorm()
```

#I don't understand the usefulness of Fuzzy Search

```{r}
?dput
```

Challenge 1
Look at the help for the c function. What kind of vector do you expect you will create if you evaluate the following:

c(1, 2, 3)
c('d', 'e', 'f')
c(1, 2, 'f')`

```{r}

? c(x)

c(1:5)

? c

```

#? c : worked. ? c() : did not work. why?

Challenge 2

Look at the help for the paste function. You’ll need to use this later. What is the difference between the sep and collapse arguments?

```{r}

? paste

paste(1:12)

paste(1:12, sep = ",")

paste(1:12, collapse = ",")
```

> paste(1:12)
 [1] "1"  "2"  "3"  "4"  "5"  "6"  "7"  "8"  "9"  "10" "11" "12"
 
> paste(1:12, sep = ",")
 [1] "1"  "2"  "3"  "4"  "5"  "6"  "7"  "8"  "9"  "10" "11" "12"
 
> paste(1:12, collapse = ",")
[1] "1,2,3,4,5,6,7,8,9,10,11,12"

#I guess don't include () with help
#sep = ???
#collapse = separates by given value but doesn't inclue a space

Challenge 3

Use help to find a function (and its associated parameters) that you could use to load data from a csv file in which columns are delimited with “\t” (tab) and the decimal point is a “.” (period). This check for decimal separator is important, especially if you are working with international colleagues, because different countries have different conventions for the decimal point (i.e. comma vs period). hint: use ??csv to lookup csv related functions.
```{r}
?? csv
```


