---
title: "R Lesson 11 Writing Data"
author: "Kelsey Deaton"
date: 2018-10-11
output: html_document
    
---

#Writing Data


#Saving Plots

* ggsave()
* export from Plots window
* pdf()

```{r}

gapminder = read.csv(path.expand("~/R.Lessons/data/gapminder_data.csv"))

pdf("Life_Exp_vs_time.pdf", width=12, height=4)
ggplot(data=gapminder, aes(x=year, y=lifeExp, colour=country)) +
  geom_line() +
  theme(legend.position = "none")

# You then have to make sure to turn off the pdf device!

dev.off()
```




#Challenge 1 

Rewrite your ‘pdf’ command to print a second page in the pdf, showing a facet plot (hint: use facet_grid) of the same data with one panel per continent.

```{r}
pdf("Life_Exp_vs_time.pdf", width = 12, height = 4)
p <- ggplot(data = gapminder, aes(x = year, y = lifeExp, colour = country)) +
  geom_line() +
  theme(legend.position = "none")
p
p + facet_grid(. ~continent)
dev.off()
```

#Writing Data

example

```{r}
#make a subset for Australia

subset_Aus = gapminder[gapminder$country == "Australia", ]

#write a data set for this subset

write.table(subset_Aus,
  file="gapminderAus.csv",
  sep=","
)
```

> SC lesson uses shell to look at file. I just opened it from Files window. 

By default R will wrap character vectors with quotation marks when writing out to file. It will also write out the row and column names. (ex: "Oceania")

Let’s fix this:

```{r}
write.table(
  gapminder[gapminder$country == "Australia",],
  file="gapminderAus.csv",
  sep=",", quote=FALSE, row.names=FALSE
)
```


Fixed. 

#Challenge 2

Write a data-cleaning script file that subsets the gapminder data to include only data points collected since 1990.

Use this script to write out the new subset to a file in the cleaned-data/ directory.

(I'll use my data folder instead)

```{r}
write.table(
  gapminder[gapminder$year >= 1990, ],
  file="~/R.Lessons/data/gapminder1990plus.csv",
  sep=",", quote=FALSE, row.names=FALSE
)
```

Wrote on my own and it Worked!


