---
title: "R lesson 13 Dataframe Manipulation with dplyr"
author: "Kelsey Deaton"
date: 2018-10-31
output:
  html_document:
    df_print: paged
---

Load data:

```{r}
gapminder = read.csv("https://raw.githubusercontent.com/swcarpentry/r-novice-gapminder/gh-pages/_episodes_rmd/data/gapminder_data.csv")
```


#dplyr

```{r}
install.packages("dplyr")
library("dplyr")
```

#Select()

select(dataframe, a, c) will return a dataframe with columns a and c

can use pipe operator %>% , which is kinda like "then"

>for more info see: https://www.datacamp.com/community/tutorials/pipe-r-tutorial#intro

in:
year_country_gdp <- gapminder %>% select(year,country,gdpPercap)

... you take the gapminder data, THEN, select the subset of data

#Filter()

Can filter a dataset to make a subset that matches the filter criteria

```{r}
year_country_gdp_euro <- gapminder %>%
    filter(continent=="Europe") %>%
    select(year,country,gdpPercap)
```

#Challenge 1

Write a single command (which can span multiple lines and includes pipes) that will produce a dataframe that has the African values for lifeExp, country and year, but not for other Continents. How many rows does your dataframe have and why?

```{r}
challenge_1 = gapminder %>%
  filter(continent=="Africa") %>%
  select(year, country, lifeExp)

nrow(challenge_1)
```

#Group_by()

Makes a subset dataframe for each of the variables that they are being grouped by. 

```{r}
gapminder %>% group_by(continent)
```
This would make a subset data frame for each continent.

#Summarize()

Performs a function (defined within summarize) on each subset data frame, then brings them back together in a summary dataframe. 

```{r}
gdp_bycontinents <- gapminder %>%
    group_by(continent) %>%
    summarize(mean_gdpPercap=mean(gdpPercap))
gdp_bycontinents
```

#Challenge 2

Calculate the average life expectancy per country. Which has the longest average life expectancy and which has the shortest average life expectancy?

```{r}
lifeExp_byCountries = gapminder %>%
  group_by(country) %>%
  summarize(mean_lifeExp = mean(lifeExp))
lifeExp_byCountries
```

To only show the shortest and longest, should have added this to above code:

```{r}
lifeExp_byCountries %>%
  filter(mean_lifeExp == min(mean_lifeExp) | mean_lifeExp == max(mean_lifeExp))
lifeExp_byCountries
```
Still get all countries when "LifeExp_byCountries" is included at the end, must no longer be filtered. 


```{r}
lifeExp_byCountries = gapminder %>%
  group_by(country) %>%
  summarize(mean_lifeExp = mean(lifeExp))
lifeExp_byCountries %>%
  filter(mean_lifeExp == min(mean_lifeExp) | mean_lifeExp == max(mean_lifeExp))

```


Can group by multiple variables.

```{r}
gdp_bycontinents_byyear <- gapminder %>%
    group_by(continent,year) %>%
    summarize(mean_gdpPercap=mean(gdpPercap))
gdp_bycontinents_byyear
```

#count() & n()

Can count the number of observations of a certain variable in a subset. 
```{r}
gapminder %>%
    filter(year == 2002) %>%
    count(continent, sort = TRUE)
```

Can have multiple summary operations. 

```{r}
gapminder %>%
    group_by(continent) %>%
    summarize(
      mean_le = mean(lifeExp),
      min_le = min(lifeExp),
      max_le = max(lifeExp),
      se_le = sd(lifeExp)/sqrt(n()))
```

#mutate()

create new variables before summarizing. 

#ifelse()

ifelse(function, do if true, do if false)

#advanced challenge

Calculate the average life expectancy in 2002 of 2 randomly selected countries for each continent. Then arrange the continent names in reverse order. Hint: Use the dplyr functions arrange() and sample_n(), they have similar syntax to other dplyr functions.

```{r}
gapminder %>%
  sample_n(2) %>%
  arrange(desc(country))
```

Worked but forgot about the "in 2002" and "calculate average lifeExp" part...

Try again:

```{r}
gapminder %>%
  group_by(country) %>%
  filter(year==2002) %>%
  summarize(meanlifeExp=mean(lifeExp)) %>%
  sample_n(2) %>%
  arrange(desc(country))

```

>SWC did way different order of operations, does it make a difference here? how? 