---
title: "R lesson 14 tidyr"
author: "Kelsey Deaton"
date: '2018-10-31'
output:
  html_document:
    df_print: paged
---
#Data frame manipulation with tidyr

load data first

```{r}
gapminder = read.csv("https://raw.githubusercontent.com/swcarpentry/r-novice-gapminder/gh-pages/_episodes_rmd/data/gapminder_data.csv")
```

```{r}
library("tidyr")
library("dplyr")
```

Wide v long dataframe format

#gather()

```{r}
gap_wide = read.csv("https://raw.githubusercontent.com/swcarpentry/r-novice-gapminder/gh-pages/_episodes_rmd/data/gapminder_wide.csv", stringsAsFactors = FALSE)
```


```{r}
gap_long <- gap_wide %>%
    gather(obstype_year, obs_values, starts_with('pop'),
           starts_with('lifeExp'), starts_with('gdpPercap'))
str(gap_long)
```

```{r}
head(gapminder)
```

```{r}
head(gap_wide)
```

```{r}
head(gap_long)
```

```{r}
gap_long <- gap_long %>% separate(obstype_year,into=c('obs_type','year'),sep="_")
gap_long$year <- as.integer(gap_long$year)
```


#Challenge 2

Using gap_long, calculate the mean life expectancy, population, and gdpPercap for each continent. Hint: use the group_by() and summarize() functions we learned in the dplyr lesson

```{r}
challenge_2 = gap_long %>%
  group_by(continent, obs_type) %>%
  summarize(mean=mean(obs_values))
challenge_2
```


#Spread()

goes from long format to intermediate

```{r}
gap_normal = gap_long %>% 
  spread(obs_type, obs_values)
gap_normal
```

#convert to wide

```{r}
gap_temp <- gap_long %>% unite(var_ID,continent,country,sep="_")
head(gap_temp)
```

> I don't understand why you would do this. 

```{r}
gap_temp <- gap_long %>%
    unite(ID_var,continent,country,sep="_") %>%
    unite(var_names,obs_type,year,sep="_")
head(gap_temp)
```

```{r}
gap_wide_new <- gap_long %>%
    unite(ID_var,continent,country,sep="_") %>%
    unite(var_names,obs_type,year,sep="_") %>%
    spread(var_names,obs_values)
head(gap_wide_new)
```

#challenge 3

Take this 1 step further and create a gap_ludicrously_wide format data by spreading over countries, year and the 3 metrics? Hint this new dataframe should only have 5 rows.

```{r}
gap_very_wide = gap_wide_new <- gap_long %>%
    unite(ID_var,continent,country,sep="_") %>%  #change should go here, by uniting obstype,year,country... then spread
    unite(var_names,obs_type,year,sep="_") %>%
    spread(var_names,obs_values) %>%
    
  

```


Answer:
```{r}
gap_ludicrously_wide <- gap_long %>%
   unite(var_names,obs_type,year,country,sep="_") %>%
   spread(var_names,obs_values)
head(gap_ludicrously_wide)
```

