---
title: "R Lesson 9 Vectorization"
author: Kelsey Deaton
date: 2018-10-01
output: html_notebook
---

R code is mostly automatically vectorized. 

#Challenge 1

Let’s try this on the pop column of the gapminder dataset.

Make a new column in the gapminder data frame that contains population in units of millions of people. Check the head or tail of the data frame to make sure it worked.

Sidenote: pop is currently reported as the full number, e.g. 1,000,000 vs 1 million.

```{r}
pop_as_million = round(gapminder$pop / 1000000, digits=1)
gapminder = cbind(gapminder, pop_as_million)
head(gapminder)
```

Looks good to me. 

#Challenge 2

On a single graph, plot population, in millions, against year, for all countries. Don’t worry about identifying which country is which.

Repeat the exercise, graphing only for China, India, and Indonesia. Again, don’t worry about which is which.

Attempt 1:
```{r}
ggplot(data=gapminder, aes(x=year, y=pop_as_million))
```
Didn't work. Need to define what type of graph!

Answer:

```{r}
ggplot(gapminder, aes(x = year, y = pop_as_million)) + geom_point #didn't have brackets!
```
not working for some reason... didn't have brackets for geom_point()

Try directly copying SC code...

```{r}
ggplot(gapminder, aes(x = year, y = pop_as_million)) +
 geom_point()
```

part b:

```{r}
subset_countries <- c("China", "India", "Indonesia")

ggplot(gapminder[gapminder$country %in% subset_countries,], 
       aes(x = year, y = pop_as_million)) +
  geom_point()
```

##Why am I still getting all countries?

