---
title: "R. Lesson 6 Week 2 Subsetting Data"
author: Kelsey Deaton
date: 2018-09-27
output: html_notebook
---

First, let's create a matrix. 

```{r}
x <- c(5.4, 6.2, 7.1, 4.8, 7.5)
names(x) <- c('a', 'b', 'c', 'd', 'e')
x
```

#Accessing elements using indices

```{r}
x[1]
x[4]
```

To ask for multiple elements, need to use c()

```{r}
x[c(1,4)]
```
to leave out c() doesn't work.
```{r}
x[1,4]
```

#Challenge 1

Given the following code:

x <- c(5.4, 6.2, 7.1, 4.8, 7.5)
names(x) <- c('a', 'b', 'c', 'd', 'e')
print(x)

output:
  a   b   c   d   e 
5.4 6.2 7.1 4.8 7.5 

Come up with at least 2 different commands that will produce the following output:

  b   c   d 
6.2 7.1 4.8 

After you find 2 d
ifferent commands, compare notes with your neighbour. Did you have different strategies?

Attempt 1
```{r}
y = c(6.2, 7.1, 4.8)
names(y) = c('b', 'c', 'd')
print(y)
```

Attempt 2
```{r}
numbers = c(6.2, 7.1, 4.8)
titles = c("b","c","d")
matrix = matrix(c(titles, numbers), nrow=2, byrow=TRUE)
matrix 
```

Worked, more or less, but not what they were going for. Use previous matrix (x) with (-)

Other option

```{r}
x[-c(1,5)]
```

#Subsetting by name or logical operations

```{r}
x[c("a", "e")]
```

#Challenge 2

Given the following code:

x <- c(5.4, 6.2, 7.1, 4.8, 7.5)
names(x) <- c('a', 'b', 'c', 'd', 'e')
print(x)

  a   b   c   d   e 
5.4 6.2 7.1 4.8 7.5 

Write a subsetting command to return the values in x that are greater than 4 and less than 7.

```{r}
x[c(x > 4, x < 7)]
```

Why didn't this work? Where are NA's from?

Instead, answer created a subset. 

```{r}
x_subset <- x[x<7 & x>4]
print(x_subset)
```

#Challenge 3 

Selecting elements of a vector that match any of a list of components is a very common data analysis task. For example, the gapminder data set contains country and continent variables, but no information between these two scales. Suppose we want to pull out information from southeast Asia: how do we set up an operation to produce a logical vector that is TRUE for all of the countries in southeast Asia and FALSE otherwise?

Suppose you have these data:

seAsia <- c("Myanmar","Thailand","Cambodia","Vietnam","Laos")
## read in the gapminder data that we downloaded in episode 2
gapminder <- read.csv("data/gapminder_data.csv", header=TRUE)
## extract the `country` column from a data frame (we'll see this later);
## convert from a factor to a character;
## and get just the non-repeated elements
countries <- unique(as.character(gapminder$country))

There’s a wrong way (using only ==), which will give you a warning; a clunky way (using the logical operators == and |); and an elegant way (using %in%). See whether you can come up with all three and explain how they (don’t) work.

gapminder data looks fine now, must have fixed when I reuploaded it. 

```{r}
seAsia <- c("Myanmar","Thailand","Cambodia","Vietnam","Laos")
## read in the gapminder data that we downloaded in episode 2

## extract the `country` column from a data frame (we'll see this later);
## convert from a factor to a character;
## and get just the non-repeated elements
countries <- unique(as.character(gapminder$country))
subset1 = countries %in% seAsia
countries
```

Not sure what I should have gotten here. (?)

#Challenge 4

Given the following code:

m <- matrix(1:18, nrow=3, ncol=6)
print(m)

     [,1] [,2] [,3] [,4] [,5] [,6]
[1,]    1    4    7   10   13   16
[2,]    2    5    8   11   14   17
[3,]    3    6    9   12   15   18

Which of the following commands will extract the values 11 and 14?

A. m[2,4,2,5]

B. m[2:5]

C. m[4:5,2]

D. m[2,c(4,5)]

I choose D. Want row 2, combined columns 4 and 5. 

Correct!


#Challenge 5

Given the following list:

xlist <- list(a = "Software Carpentry", b = 1:10, data = head(iris))

Using your knowledge of both list and vector subsetting, extract the number 2 from xlist. Hint: the number 2 is contained within the “b” item in the list.

Attempt 1:
```{r}

xlist <- list(a = "Software Carpentry", b = 1:10, data = head(iris))

xlist[[b[2]]]
```

Attempt 2:
```{r}
xlist[[2[2]]]
```

No again. Answer: 

```{r}
xlist$b[2]
```

or:
This makes more sense:
```{r}
xlist[[2]][2]
```

#Challenge 6

Given a linear model:

```{r}
mod <- aov(pop ~ lifeExp, data=gapminder)
? aov()
```

Extract the residual degrees of freedom (hint: attributes() will help you)

Didn't attempt since having an issue with gapminder data again. 

#Data frames

```{r}
head(gapminder[3])
```

#How to properly import data!

Can keep the ~ if you use path.expand, actually useful in case your home directory or computer changes. 

Need to use "" to make into a string. 

```{r}
gapminder = read.csv(path.expand("~/R.Lessons/data/gapminder_data.csv"))
```

#Retry Challenge 6

Given a linear model:

```{r}
mod <- aov(pop ~ lifeExp, data=gapminder)
? aov()
```

FYI: aov is ANOVA. (dependent ~ independent variable)

Extract the residual degrees of freedom (hint: attributes() will help you)

```{r}
? attributes
attributes(mod)
attributes(mod$df.residual)
mod$df.residual
```
Used last command "mod$df.residual" to find the number

#Challenge 7

Fix each of these errors.

1. Extract observations collected for the year 1957

With error:
```{r}
gapminder[gapminder$year = 1957,]
```

Fixed by changing "=" to "=="

```{r}
gapminder[gapminder$year == 1957,]
```

2. Extract all columns except 1 through to 4

With error:
```{r}
gapminder[,-1:4]
```

Fixed: need to -c()

```{r}
gapminder[,-c(1:4)]
```


3. Extract the rows where the life expectancy is longer the 80 years

With error:
```{r}
gapminder[gapminder$lifeExp > 80]
```

Fixed: missing a comma to distinguish between rows and column. 

```{r}
gapminder[gapminder$lifeExp > 80, ]
```


4. Extract the first row, and the fourth and fifth columns (lifeExp and gdpPercap).

```{r}
gapminder[1, 4, 5]
```

Interesting that it didn't give an error. Not sure what answer it gave instead. 

Fixed: (not)

```{r}
gapminder[1, 4:5]
```

Correct. 

Also correct: (this was SC's answer:)

```{r}
gapminder[1, c(4, 5)]
```



5. Advanced: extract rows that contain information for the years 2002 and 2007

With error:
```{r}
gapminder[gapminder$year == 2002 | 2007,]
```

Attempt: works!

```{r}
gapminder[c(gapminder$year == 2002, gapminder$year == 2002),]
```

SC answer 1: 

```{r}
gapminder[gapminder$year == 2002 | gapminder$year == 2007,]
 
```

SC answer 2:

```{r}
gapminder[gapminder$year %in% c(2002, 2007),]
```

#Challenge 8

Why does gapminder[1:20] return an error? How does it differ from gapminder[1:20, ]?



```{r}
gapminder[1:20]
```
Leaves out columns.

```{r}
gapminder[1:20, ]
```
Includes "," to specify rows, columns. 

Create a new data.frame called gapminder_small that only contains rows 1 through 9 and 19 through 23. You can do this in one or two steps.

```{r}
gapminder_small = c(gapminder[1:9,], gapminder[19:23,])
gapminder_small
```
Why did this come out as lists?

SC's better answer: 

```{r}
gapminder_small_sc = gapminder[c(1:9, 19:23),]
gapminder_small_sc
```

