---
title: "R Lesson 5 Exploring Data Frames"
author: Kelsey Deaton
date: 2018-09-21
output: html_notebook
---

Run lesson 4 so that history is loaded, because we're using the same "cats" matrix. 

#adding columns and rows

```{r}
age <- c(2, 3, 5)
cats
```

add new age vector to cats matrix. need to have same number of entries as rows.

```{r}
cbind(cats, age)
```

```{r}
nrow(cats)

length(age)
```

```{r}
cats$coat
```

```{r}
levels(cats$coat)
```

would get an error if you tried to add another row without modifying the factor first
to add another "level" to the factor "coat type" 

```{r}
levels(cats$coat) = c(levels(cats$coat), "tortoiseshell")
levels(cats$coat)
```

now its added. 

#now you can add another row to the matrix. 

```{r}
cats = rbind(cats, list("tortoiseshell", 3.3, TRUE, 9))
cats
```

#change factor into character

```{r}
str(cats)
```

Age wasn't kept...

```{r}
ncol(cats)
```

```{r}
cats
```

Try redefining cats with age. 

```{r}
cats = cbind(cats, age)
```

Age didn't include the 4th tortoiseshell cat. so redefine age. 

```{r}
age = c(2,3,5,9)
age
```

Now retry redefining cats. 

```{r}
cats = cbind(cats, age)
cats
```

worked!

OK Back to ...

#change factor into character

```{r}
str(cats)
```

```{r}
cats$coat <- as.character(cats$coat)
str(cats)
```

#Challenge 1

Let’s imagine that 1 human year is equivalent to 7 cat years.

Create a vector called human_age by multiplying cats$age by 7.
Convert human_age to a factor.
Convert human_age back to a numeric vector using the as.numeric() function. Now divide it by 7 to get the original ages back. Explain what happened.


```{r}
human_age = cats$age * 7
human_age
```

```{r}
human_age = as.factor(human_age)
typeof(human_age)
```

```{r}
human_age = as.numeric(human_age)
typeof(human_age)
human_age
```

#important!

value of the ages has changed. The ages were changed to factors, which are given a level of 1:4. Converting back to numeric gives the value of the level, not the original age. Need to convert back to character first. 

#Removing rows

Current cats matrix. 

```{r}
cats
```

Remove 3rd row. 

```{r}
cats[-3, ]
```

Can remove any rows with an NA with command "na.omit(cats)"

#Remove columns

```{r}
cats[ ,-3]
```

Changes were not permanent though. Would have to reassign the changes to "cats"

```{r}
cats
```


#Challenge 2

You can create a new data frame right from within R with the following syntax:

df <- data.frame(id = c("a", "b", "c"),
                 x = 1:3,
                 y = c(TRUE, TRUE, FALSE),
                 stringsAsFactors = FALSE)
Make a data frame that holds the following information for yourself:

first name
last name
lucky number
Then use rbind to add an entry for the people sitting beside you. Finally, use cbind to add a column with each person’s answer to the question, “Is it time for coffee break?”


```{r}
ked = data.frame(id = c("Kelsey"),
                  lucky_number = 28,
                  coffee_break = TRUE,
                  stringsAsFactors = FALSE)
ked
```

```{r}
tnw = data.frame(id = c("Talia"),
                        lucky_number = 7,
                        coffee_break = TRUE, 
                        stringsAsFactors = FALSE)
tnw
```

```{r}
both = rbind(ked, tnw)
both
```

Add last names.

```{r}
both = cbind(both, lastnames = c("Deaton", "Neal"))
both
```


