---
title: "R Lesson 12 Data Frames with Plyr"
author: "Kelsey Deaton"
date: 2018-10-12
output: html_document
---

Load gapminder data

```{r}
gapminder = read.csv(path.expand("~/R.Lessons/data/gapminder_data.csv"))

```

From lesson 11:
```{r}
# Takes a dataset and multiplies the population column
# with the GDP per capita column.
calcGDP <- function(dat, year=NULL, country=NULL) {
  if(!is.null(year)) {
    dat <- dat[dat$year %in% year, ]
  }
  if (!is.null(country)) {
    dat <- dat[dat$country %in% country,]
  }
  gdp <- dat$pop * dat$gdpPercap

  new <- cbind(dat, gdp=gdp)
  return(new)
}
```

what if we wanted to calculated the mean GDP per continent?
We could run calcGDP and then take the mean of each continent:

```{r}
withGDP <- calcGDP(gapminder)
mean(withGDP[withGDP$continent == "Africa", "gdp"])
```

But this isn't very "nice"

Instead We want to split our data into groups, in this case continents, apply some calculations on that group, then optionally combine the results together afterwards.

#plyr package

```{r}
library("plyr")
```

xxply(.data, .variables, .function)
where x = "a" = array, "l" = list, "d" = data frame, "r" = n replicates, "m" = function arguments

example:

```{r}
ddply(
 .data = calcGDP(gapminder),
 .variables = "continent",
 .fun = function(x) mean(x$gdp)
)
```

Calculate the average life expectancy per continent. Which has the longest? Which had the shortest?

```{r}
ddply(
 .data = gapminder,
 .variables = "continent",
 .fun = function(x) mean(x$lifeExp)
)
```

#Challenge 2

Calculate the average life expectancy per continent and year. Which had the longest and shortest in 2007? Which had the greatest change in between 1952 and 2007?

```{r}
ddply(
  .data = gapminder,
  .variables = c("continent", "year"),
  .fun = function(x) mean(x$lifeExp)
)
```


