---
title: "R Notebook"
author: "Kelsey Deaton"
date: 2018-10-10
output: html_document

---

#Making your own functions

```{r}
fahr_to_kelvin <- function(temp) {
  kelvin <- ((temp - 32) * (5 / 9)) + 273.15
  return(kelvin)
}

fahr_to_kelvin(32)
```


#Challenge 1

Write a function called kelvin_to_celsius() that takes a temperature in Kelvin and returns that temperature in Celsius.

Hint: To convert from Kelvin to Celsius you subtract 273.15

```{r}
kelvin_to_celsius = function(temp) {
  celsius = temp - 273.15
  return(celsius)
}

kelvin_to_celsius(273.15)
```

check!

#Combining Functions

Define the function to convert directly from Fahrenheit to Celsius, by reusing the two functions above (or using your own functions if you prefer).

```{r}
fahr_to_celsius = function(temp_f) {
  temp_k = fahr_to_kelvin(temp_f)
  temp_c = kelvin_to_celsius(temp_k)
  return(temp_c)
}

fahr_to_celsius(32)
```

#Defensive Programming

Can include conditional statements so that function only runs if conditions are met. 

```{r}
fahr_to_kelvin <- function(temp) {
  if (!is.numeric(temp)) {
    stop("temp must be a numeric vector.") #stops function if temp is not numeric
  }
  kelvin <- ((temp - 32) * (5 / 9)) + 273.15
  return(kelvin)
}
```

Can use function stopifnot()

```{r}
fahr_to_kelvin <- function(temp) {
  stopifnot(is.numeric(temp))
  kelvin <- ((temp - 32) * (5 / 9)) + 273.15
  return(kelvin)
}
```

Let's check: 

```{r}
fahr_to_kelvin(32)
```

```{r}
#fahr_to_kelvin(hot) gives an error: Error in stopifnot(is.numeric(temp)) : object 'hot' not found
```

Works. 

#Challenge 3

Exactly same code as example above. 

#More combining functions

Make a function to calculate gdp per capita

```{r}
gapminder = read.csv(path.expand("~/R.Lessons/data/gapminder_data.csv"))

# Takes a dataset and multiplies the population column
# with the GDP per capita column.
calcGDP <- function(dat) {
  gdp <- dat$pop * dat$gdpPercap
  return(gdp)
}

calcGDP(head(gapminder))
```


 so there’s a lot going on in this function now. In plain English, the function now subsets the provided data by year if the year argument isn’t empty, then subsets the result by country if the country argument isn’t empty. Then it calculates the GDP for whatever subset emerges from the previous two steps. The function then adds the GDP as a new column to the subsetted data and returns this as the final result. You can see that the output is much more informative than a vector of numbers:
```{r}
# Takes a dataset and multiplies the population column
# with the GDP per capita column.
calcGDP <- function(dat, year=NULL, country=NULL) {  #Here we’ve added two arguments, year, and country. We’ve set default arguments for both as NULL using the = operator in the function definition. This means that those arguments will take on those values unless the user specifies otherwise.
  if(!is.null(year)) {
    dat <- dat[dat$year %in% year, ]
  }
  if (!is.null(country)) {
    dat <- dat[dat$country %in% country,]
  }
  gdp <- dat$pop * dat$gdpPercap

  new <- cbind(dat, gdp=gdp)
  return(new)
}

head(calcGDP(gapminder, year=2007))
```

#Challenge 4

Test out your GDP function by calculating the GDP for New Zealand in 1987. How does this differ from New Zealand’s GDP in 1952?

```{r}
calcGDP(gapminder, year=1987, country="New Zealand")
```

#Challenge 5

The paste() function can be used to combine text together, e.g:

best_practice <- c("Write", "programs", "for", "people", "not", "computers")
paste(best_practice, collapse=" ")
[1] "Write programs for people not computers"
Write a function called fence() that takes two vectors as arguments, called text and wrapper, and prints out the text wrapped with the wrapper:

fence(text=best_practice, wrapper="***")
Note: the paste() function has an argument called sep, which specifies the separator between text. The default is a space: “ “. The default for paste0() is no space “”.

```{r}
fence <- function(text, wrapper){
  text <- c(wrapper, text, wrapper)
  result <- paste(text, collapse = " ")
  return(result)
}
best_practice <- c("Write", "programs", "for", "people", "not", "computers")
fence(text=best_practice, wrapper="***")
```

