---
title: "R Lesson 8 Creating Graphics"
author: Kelsey Deaton
date: 2018-09-30
output: html_notebook
---

Creating graphics using ggplot2

#Challenge 1

Challenge 1
Modify the example so that the figure shows how life expectancy has changed over time:

Hint: the gapminder dataset has a column called “year”, which should appear on the x-axis.

SC Example:
```{r}
library("ggplot2")

ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp)) + geom_point()
```

Modified: just change x = gdppercap to x = year

```{r}
ggplot(data = gapminder, aes(x = year, y = lifeExp)) + geom_point()
```

#Challenge 2

In the previous examples and challenge we’ve used the aes function to tell the scatterplot geom about the x and y locations of each point. Another aesthetic property we can modify is the point color. Modify the code from the previous challenge to color the points by the “continent” column. What trends do you see in the data? Are they what you expected?

```{r}

ggplot(data = gapminder, aes(x = year, y = lifeExp, color = continent)) + geom_point()

```

#Layers

