---
title: "R Lesson 9 Vectorization"
author: "Kelsey Deaton"
date: '2018-10-01'
output: html_document
    
---

R code is mostly automatically vectorized. 

#Challenge 1

Let’s try this on the pop column of the gapminder dataset.

Make a new column in the gapminder data frame that contains population in units of millions of people. Check the head or tail of the data frame to make sure it worked.

Sidenote: pop is currently reported as the full number, e.g. 1,000,000 vs 1 million.

```{r}
gapminder = read.csv(path.expand("~/R.Lessons/data/gapminder_data.csv"))
pop_as_million = round(gapminder$pop / 1000000, digits=1)
gapminder = cbind(gapminder, pop_as_million)
head(gapminder)
```

Looks good to me. 

#Challenge 2

On a single graph, plot population, in millions, against year, for all countries. Don’t worry about identifying which country is which.

Repeat the exercise, graphing only for China, India, and Indonesia. Again, don’t worry about which is which.

Attempt 1:
```{r}
ggplot(data=gapminder, aes(x=year, y=pop_as_million))
```
Didn't work. Need to define what type of graph!

Answer:

```{r}
#ggplot(gapminder, aes(x = year, y = pop_as_million)) + geom_point #didn't have brackets!
```
Was not working... didn't have brackets for geom_point()

Try directly copying SC code...

```{r}
ggplot(gapminder, aes(x = year, y = pop_as_million)) +
 geom_point()
```

part b:

```{r}
subset_countries <- c("China", "India", "Indonesia")

ggplot(gapminder[gapminder$country %in% subset_countries,], 
       aes(x = year, y = pop_as_million)) +
  geom_point()
```

#Challenge 3

Given the following matrix:

m <- matrix(1:12, nrow=3, ncol=4)
m
     [,1] [,2] [,3] [,4]
[1,]    1    4    7   10
[2,]    2    5    8   11
[3,]    3    6    9   12
Write down what you think will happen when you run:

1) m ^ -1
2) m * c(1, 0, -1)
3) m > c(0, 20)
4) m * c(1, 0, -1, 2)

Answers
1) Each number in the matrix will be raised ^-1
2)... I don't know they didn't go over that. 

```{r}
m <- matrix(1:12, nrow=3, ncol=4)
m
```

1)
```{r}
m ^ -1
```

2)
```{r}
m * c(1, 0, -1)
```

Result: first row was multiplied by first value, 2nd row by 2nd value, etc

3) guess = logical, first and third row will be compared to first value (0) = FALSE
while the second row will be compared to the second value = TRUE

```{r}
m > c(0, 20)
```

Result: my guess was wrong, which makes me think my guess for the reason for the result for #2 was wrong as well. The operation goes by each element in order, which happened to also correspond to the rows for #2. 

4)  m * c(1, 0, -1, 2)
my guess = multiplies 1st element by 1, 2nd by 0, etc corresponding to the elements in rows of first column, then second, etc

```{r}
m * c(1, 0, -1, 2)
```
Yes

#Challenge 4

We’re interested in looking at the sum of the following sequence of fractions:

 x = 1/(1^2) + 1/(2^2) + 1/(3^2) + ... + 1/(n^2)
This would be tedious to type out, and impossible for high values of n. Use vectorisation to compute x when n=100. What is the sum when n=10,000?

```{r}
sum(1/(1:100)^2)
```

```{r}
sum(1/(1:10000)^2)
```

