---
title: "R lesson 4 Data Structures"
author: Kelsey Deaton
date: 9/17/2018
output: html_notebook
---
Why isn't the html showing any of the following?
Maybe because I leave all my wrong code? -> try putting wrong code as #
#no, still doens't work

Changed output from html to html_notebook

```{r}
cats <- read.csv(file = "/home/guest/R.Lessons/data/feline_data.csv")
cats

```

#the csv file didn't work with a header "R lesson 4" at the top, needed to be in standard csv function

#print out has extra html symbols...

```{r}
cats$weight
```

```{r}
cats$coat
```

#```{r}
#file.show("data/feline_data.csv")
#```
```{r}
file.show("/home/guest/R.Lessons/data/feline_data.csv")
```

```{r}
coats <- c('tabby', 'tortoiseshell', 'tortoiseshell', 'black', 'tabby')
coats
str(coats)
categories <- factor(coats)
class(categories)
```


###Challenges

Challenge 1: Start by making a vector with the numbers 1 through 26. Multiply the vector by 2, and give the resulting vector names A through Z (hint: there is a built in vector called LETTERS)

```{r}
vector <- 1:26
vector

```

```{r}
vector*2
```

```{r}
names(vector) <- LETTERS
names(vector)
```

#Challenge 2: Is there a factor in our cats data.frame? what is its name? Try using ?read.csv to figure out how to keep text columns as character vectors instead of factors; then write a command or two to show that the factor in cats is actually a character vector when loaded in this way.

```{r}

```

