---
title: "R Lesson 5 Exploring Data Frames"
author: Kelsey Deaton
date: 2018-09-21
output: html_notebook
---

Run lesson 4 so that history is loaded, because we're using the same "cats" matrix. 

#adding columns and rows

```{r}
age <- c(2, 3, 5)
cats
```

add new age vector to cats matrix. need to have same number of entries as rows.

```{r}
cbind(cats, age)
```

```{r}
nrow(cats)

length(age)
```

```{r}
cats$coat
```

```{r}
levels(cats$coat)
```

would get an error if you tried to add another row without modifying the factor first
to add another "level" to the factor "coat type" 

```{r}
levels(cats$coat) = c(levels(cats$coat), "tortoiseshell")
levels(cats$coat)
```

now its added. 

#now you can add another row to the matrix. 

```{r}
cats = rbind(cats, list("tortoiseshell", 3.3, TRUE, 9))
cats
```

#change factor into character

```{r}
str(cats)
```

Age wasn't kept...

```{r}
ncol(cats)
```

```{r}
cats
```

Try redefining cats with age. 

```{r}
cats = cbind(cats, age)
```

Age didn't include the 4th tortoiseshell cat. so redefine age. 

```{r}
age = c(2,3,5,9)
age
```

Now retry redefining cats. 

```{r}
cats = cbind(cats, age)
cats
```

worked!

OK Back to ...

#change factor into character

```{r}
str(cats)
```

```{r}
cats$coat <- as.character(cats$coat)
str(cats)
```

#Challenge 1

Let’s imagine that 1 human year is equivalent to 7 cat years.

Create a vector called human_age by multiplying cats$age by 7.
Convert human_age to a factor.
Convert human_age back to a numeric vector using the as.numeric() function. Now divide it by 7 to get the original ages back. Explain what happened.


```{r}
human_age = cats$age * 7
human_age
```

```{r}
human_age = as.factor(human_age)
typeof(human_age)
```

```{r}
human_age = as.numeric(human_age)
typeof(human_age)
human_age
```

#important!

value of the ages has changed. The ages were changed to factors, which are given a level of 1:4. Converting back to numeric gives the value of the level, not the original age. Need to convert back to character first. 

#Removing rows

Current cats matrix. 

```{r}
cats
```

Remove 3rd row. 

```{r}
cats[-3, ]
```

Can remove any rows with an NA with command "na.omit(cats)"

#Remove columns

```{r}
cats[ ,-3]
```

Changes were not permanent though. Would have to reassign the changes to "cats"

```{r}
cats
```


#Challenge 2

You can create a new data frame right from within R with the following syntax:

df <- data.frame(id = c("a", "b", "c"),
                 x = 1:3,
                 y = c(TRUE, TRUE, FALSE),
                 stringsAsFactors = FALSE)
Make a data frame that holds the following information for yourself:

first name
last name
lucky number
Then use rbind to add an entry for the people sitting beside you. Finally, use cbind to add a column with each person’s answer to the question, “Is it time for coffee break?”


```{r}
ked = data.frame(id = c("Kelsey"),
                  lucky_number = 28,
                  coffee_break = TRUE,
                  stringsAsFactors = FALSE)
ked
```

```{r}
tnw = data.frame(id = c("Talia"),
                        lucky_number = 7,
                        coffee_break = TRUE, 
                        stringsAsFactors = FALSE)
tnw
```

```{r}
both = rbind(ked, tnw)
both
```

Add last names.

```{r}
both = cbind(both, lastnames = c("Deaton", "Neal"))
both
```

#Realistic Example

```{r}
gapminder = read.csv("~/R.Lessons/data/gapminder_data.csv")
```

Went over to Files and clicked on "gapminder_data" then clicked "Import Dataset"

```{r}
str(gapminder)
```

Looks like I have different data... SF has "
'data.frame':	1704 obs. of  6 variables:
 $ country  : Factor w/ 142 levels "Afghanistan",..: 1 1 1 1 1 1 1 1 1 1 ...
 $ year     : int  1952 1957 1962 1967 1972 1977 1982 1987 1992 1997 ...
 $ pop      : num  8425333 9240934 10267083 11537966 13079460 ...
 $ continent: Factor w/ 5 levels "Africa","Americas",..: 3 3 3 3 3 3 3 3 3 3 ...
 $ lifeExp  : num  28.8 30.3 32 34 36.1 ...
 $ gdpPercap: num  779 821 853 836 740 ...
  "
  
```{r}
typeof(gapminder$year)
```

```{r}
typeof(gapminder$country)
```

#Challenge 3

It’s good practice to also check the last few lines of your data and some in the middle. How would you do this?

Searching for ones specifically in the middle isn’t too hard but we could simply ask for a few lines at random. How would you code this?

Tail:
```{r}
tail(gapminder)
```

#Challenge 4

Go to file -> new file -> R script, and write an R script to load in the gapminder dataset. Put it in the scripts/ directory and add it to version control.

Run the script using the source function, using the file path as its argument (or by pressing the “source” button in RStudio).

Didn't work.


